Bayesian Hyper-Heuristics

### 2023-02-01 article/sections/02-artificial-neural-networks.tex
🚧 (article): Save point, WIP

>>>>

\acp{ANN} can generally be described as well-organised structures of mathematical computation and are inspired by the biological brain  \cite{ref:engelbrecht:2007}. \acp{ANN} have been successfully applied to a range of problem classes.  \citeauthor{ref:engelbrecht:2007}  \cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN}. These include classification  \cite{ref:khan:2001}, pattern matching  \cite{ref:cannady:1998, ref:kumar:1994}, pattern completion  \cite{ref:dayhoff:2001}, optimisation  \cite{ref:specht:1991} and data mining  \cite{ref:singh:2009}. \Acp{AN} provide the fundamental building blocks for \acp{ANN}.

The fundamental building blocks of \acp{ANN} is the \acs{AN}. The \acs{AN} implements a non-linear mapping from some $I$-dimensional, real-numbered input space to some $T$-dimensional, real-numbered output space. The output space is usually in the ranges $[0,1]$ or $[-1,1]$, depending on the problem being solved \cite{ref:engelbrecht:2007}. In essence, the \acs{AN} implements a weighted sum of its inputs and applies some non-linear transformation as a decision making boundary.

Multiple \acp{AN} can be organised and used together forming a ``network'' of \acp{AN} referred to as an \acf{ANN}. The architecture of the \acs{ANN} refers to the way in which \acp{AN} are organised. \acp{ANN} can be organised in layers where a single layer can contain multiple \acp{AN}. Output from one layer is propagated as input to the next layer. The topology of the \acs{ANN} refers to the way that layers of \acp{AN} are connected to each other. There are many different topologies  \cite{ref:miikkulainen:2010}.

This research focuses on a particular type of \acs{ANN}, referred to as \acp{FFNN}. \acp{FFNN} were the first and simplest type of \acp{ANN} developed  \cite{ref:schmidhuber:2015} and implement an architecture consisting of input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement fully connected topologies, where each \acs{AN} in one layer is connected to all \acp{AN} in the next, without any cycles  \cite{ref:zell:1994}.In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes.

\textit{Training} is the process whereby the weights of the \acs{FFNN} are systematically changed with the aim of improving the \textit{performance} of the \acs{FFNN}. Finding the optimal weights that produce the best performance on a given task is an optimisation problem. The optimisation algorithm used to find the optimal weights is referred to as a \textit{heuristic}. Heuristics search for possible solutions in the solution-space and make use of information from the search space to guide to process.

During the training process, the \acs{FFNN} is exposed to data while trying to produce some target outcome. The degree to which the produced outcome differs from the target outcome is referred to as \textit{loss}. Since training of \acp{FFNN} is an optimisation problem, the goal of the training process is to minimise the loss. The loss is calculated using an error function.

Supervised learning is the process of training where the data that is presented to the \acs{FFNN} during training, includes the desired solution  \cite{ref:geron:2017}. The \acs{FFNN} learns the mapping function from the input to the target output  \cite{ref:brownlee:2016}.

### 2023-02-01 article/sections/03-heuristics.tex
🚧 (article): Save point, WIP

A \index{heuristic}heuristic refers to an algorithmic search technique that serves as a guide to a search process where good solutions to a optimisation problem is being sought out. Many different techniques have been used to train \acp{FFNN} \cite{ref:kingma:2014}. At the time of writing, the majority of work that is published around the training of \acp{FFNN} involves the use of gradient-based techniques \cite{ref:nel:2021}.

\subsection{Gradient-Based Heuristics}\label{sec:heuristics:gd}

Gradient-based \index{heuristic}heuristics are optimisation techniques that make use of derivates obtained from evaluating the \acs{ANN} being trained. In the context of supervised learning, loss functions produce a scalar value $\epsilon$, that represents the error between the output of the \acs{ANN} and the desired output. When using \acs{GD} to train \acp{ANN}, the \index{loss function}loss function is used to adjust the weights of the \acs{ANN} in order to minimise the error \cite{ref:engelbrecht:2007}.

The simplest type of \acs{GD} algorithm is referred to as \acs{SGD}, which implements a gradient-based weight update step for each training pattern. The weight update step and is given as

\begin{equation}
	\label{eq:heuristics:gd:sgd}
	\begin{split}
		\boldsymbol{w} = \boldsymbol{w} - \eta \boldsymbol{g}
	\end{split}
\end{equation}

where $\boldsymbol{w}$ refers to the weight vector, $\eta$ refers to the learning rate, which controls the step size and $\boldsymbol{g}$ refers to the gradient of the error function relative to weight vector.

There are many variants of gradient-based \index{heuristic}heuristics, however, they all fundamentally apply the same generic \acf{GD} framework that propagates the error signal backwards through the \acs{ANN}. This algorithm is known as \acf{BP}. \Acs{BP} was popularised by \citeauthor{ref:werbos:1994} \cite{ref:werbos:1994}. \citeauthor{ref:engelbrecht:2007} \cite{ref:engelbrecht:2007} describes the \acs{BP} process in two steps. During the forward pass step, the output values of the \acs{FFNN} is calculated for each training pattern. Then, during the backward propogation step, the error signal is propagated backwards from the output layer, through the hidden layer, to the input layer of the \acs{FFNN}. Weights at the output and hidden layers are then adjusted as functions of the backpropagated error signal.

In the context of this research, the implementation of \acs{SGD} refers to the mini-batch training implementation of \acs{GD}, where a small batch of training patterns are fed to the \acs{FFNN} at once and the error function is aggregated across all training patterns.

@@

\subsection{Meta-Heuristics}\label{sec:heuristics:mh}

Gradient-based heuristics are sensitive to the problem that they are applied to, with hyper-parameter selection often dominating the research focus \cite{ref:bengio:2000, ref:feurer:2019}. \citeauthor{ref:blum:2003} \cite{ref:blum:2003} mention that since the 1980's, a new kind of approximate algorithm has emerged which tries to combine basic heuristic methods in higher level frameworks aimed at efficiently and effectively exploring a search space. These methods are referred to as \index{meta-heuristic}\textit{meta-heuristics}.

>>>>

A heuristic refers to an algorithmic search technique that serves as a guide to a search process where good solutions to a optimisation problem is being sought out. Many different techniques have been used to train \acp{FFNN} \cite{ref:kingma:2014}. At the time of writing, the majority of work that is published around the training of \acp{FFNN} involves the use of gradient-based techniques \cite{ref:nel:2021}.

Gradient-based heuristics are optimisation techniques that make use of derivates obtained from evaluating the \acs{ANN} being trained. In the context of supervised learning, loss functions produce a scalar value $\epsilon$, that represents the error between the output of the \acs{ANN} and the desired output. When using \acs{GD} to train \acp{ANN}, the loss function is used to adjust the weights of the \acs{ANN} in order to minimise the error \cite{ref:engelbrecht:2007}.

There are many variants of gradient-based heuristics, however, they all fundamentally apply the same generic \acf{GD} framework that propagates the error signal backwards through the \acs{ANN}. This algorithm is known as \acf{BP}. \Acs{BP} was popularised by \citeauthor{ref:werbos:1994} \cite{ref:werbos:1994}.

The simplest type of \acs{GD} algorithm is referred to as \acs{SGD}, which implements a gradient-based weight update step for each training pattern. In the context of this research, the implementation of \acs{SGD} refers to the mini-batch training implementation of \acs{GD}, where a small batch of training patterns are fed to the \acs{FFNN} at once and the error function is aggregated across all training patterns.

@@

Gradient-based heuristics are sensitive to the problem that they are applied to, with hyper-parameter selection often dominating the research focus \cite{ref:bengio:2000, ref:feurer:2019}. \citeauthor{ref:blum:2003} \cite{ref:blum:2003} mention that since the 1980's, a new kind of approximate algorithm has emerged which tries to combine basic heuristic methods in higher level frameworks aimed at efficiently and effectively exploring a search space. These methods are referred to as \acp{MH}.

### 2023-02-01 article/sections/04-hyper-heuristics.tex
🚧 (article): Save point, WIP

~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} define \acp{HH} as search methods or learning mechanism for selecting or generating heuristics to solve computational search problems.~\citeauthor{ref:burke:2003}~\cite{ref:burke:2003} mentions that a \acs{HH} is a high-level approach that, given a particular problem instance and a number of low-level heuristics, can select and apply an appropriate low-level heuristic at each decision point.

\citeauthor{ref:grobler:2015}~\cite{ref:grobler:2015} states that \acp{HH} promote the design of more generally applicable search methodologies and tend to perform relatively well on a large set of different problems, in contrast to specialised algorithms, which typically focus on outperforming the state-of-the-art for a single application.

\acp{HH} implement a form of \index{meta-learning}meta-learning that is concerned with the selection of the best heuristic from a pool of heuristics to solve a given problem. In the context of population-based \acp{HH}, an entity pool exists that represent a pool of candidate solutions to the given problem. Each entity in the entity pool is assigned its own low-level \index{heuristic}heuristic from the heuristic pool.

The selection of the best heuristic to apply to a candidate solution, is based on the performance of the heuristic relative to that particular candidate solution at a particular point in the search process. It can be said that \acp{HH} are concerned with finding the best \index{heuristic}heuristic in \textit{heuristic space}, while the underlying low-level \index{heuristic}heuristics find solutions in the feasible \textit{search/solution space}.

\citeauthor{ref:grobler:2015}~\cite{ref:grobler:2015} highlights two fundamental ideas behind \acp{HH}. Firstly, the recognition that the process of selecting or designing efficient hybrid and/or cooperative heuristics can be regarded as a computational search problem in itself. Secondly, there is significant potential to improve search methodologies by the incorporation of learning mechanisms that can adaptively guide the search. These two fundamental ideas have inspired different types of hyper-heuristics~\cite{ref:burke:2010}.

In the general context of optimisation, many different types of \acp{HH} have been implemented and applied to many different problems. Some notable examples include the \index{simulated annealing}simulated annealing-based \acs{HH} by~\citeauthor{ref:dowsland:2007}~\cite{ref:dowsland:2007}, the \index{tabu-search}tabu-search \acs{HH} of~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}, the \index{heterogeneous meta-hyper-heuristic}heterogeneous meta-hyper-heuristic by~\citeauthor{ref:grobler:2012}~\cite{ref:grobler:2012} and work done by~\citeauthor{ref:vanderstockt:2018}~\cite{ref:vanderstockt:2018} on the analysis of selection hyper-heuristics for population-based \acp{MH} in real-valued dynamic optimization. Research on the application of \acp{HH} in the context of \acs{FFNN} training is still scarce.~\citeauthor{ref:nel:2021}~\cite{ref:nel:2021} provides the first research in this field, applying a \acs{HH} to \acs{FFNN} training. The following section provides a framework for \acs{HH} classification.

\subsection{Classification of Hyper-Heuristics}\label{sec:hh:classification}

\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} proposed a modern classification scheme used to classify \acp{HH}. According to the proposed classification scheme, \acp{HH} are classified in two dimensions. These include the \textit{source of feedback} used during learning and the nature of the \textit{heuristic search space}. Figure \ref{fig:heuristics:hh:classification} presents the classification scheme as proposed by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.85\textwidth]{images/hh_classification.pdf}
	\caption{A classification of \acs{HH} approaches, according to two dimensions: (i) the source of feedback used during learning, and (ii) the nature of the heuristic search space.}
	\label{fig:heuristics:hh:classification}
\end{figure}

\subsubsection{Source of Feedback}

\acp{HH} make use of feedback from the search process to adapt and guide the search process. Some \acp{HH} implement a learning mechanism that utilises this feedback. Learning \acp{HH} can implement a form of \index{online learning}\textit{online learning} or \index{offline learning}\textit{offline learning}.

\acp{HH} that implement \index{online learning}online learning, implement a form of learning that continues to takes place while the algorithm is solving an instance of the underlying optimisation problem. Task-dependent, local properties can be used by the high-level \index{heuristic}heuristic to determine the appropriate low-level heuristic to apply at various points in the search process. Examples of \index{online learning}online learning approaches within \acp{HH} include the use of \acf{RL} and \acfp{MH} as high-level search strategies.

\acp{HH} that implement \index{offline learning}offline learning, implement a form of learning where knowledge, in the form of rules or programs, is gathered from a set of training instances and is then applied independently from the search process itself. Examples of offline learning approaches within hyper-heuristics include \textit{learning classifier systems} and \textit{case-based reasoning}.

\subsubsection{Heuristic Search Space}
According to the classification proposed by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}, the second dimension of classification involves the nature of the \index{heuristic}heuristic search space. Distinction is made between \index{heuristic}heuristic \textit{selection} and heuristic \textit{generation}.

\subsubsection{Selection vs. Generation}

Selection \acp{HH} implement a high-level search mechanism that is used to determine which heuristic to apply to the underlying optimisation problem at a given point in the optimisation process. Selection mechanisms can include probabilistic approaches or high level \acfp{MH}. On the contrary, \index{heuristic}heuristic generation methodologies implement a mechanism that generate new heuristics from a pool of \textit{components} of various heuristics. Generation \acp{HH} often make use of \acfp{EA}~\cite{ref:burke:2010}.

Selection \acp{HH} consist of two components, including a low-level \index{heuristic}heuristic selection strategy and a move acceptance strategy. Low-level \index{heuristic}heuristic selection can be done in a simple, \textit{non-adaptive} way. No learning is involved in these approaches. For non-adaptive techniques, heuristic selection is based on a predefined heuristic ordering that is generated either randomly or in a ordered cycle that is repeated throughout the optimisation process~\cite{ref:cowling:2000}. As an alternative, heuristic selection may incorporate an \textit{adaptive} (\index{online learning}\textit{online learning}) mechanism, based on the probabilistic weighting of the low-level heuristics~\cite{ref:burke:2003} or some type of performance statistics~\cite{ref:cowling:2000}.

A move acceptance strategy is the mechanism by which the application of a low-level \index{heuristic}heuristic is either accepted or rejected. In general, a move is accepted or rejected based on the quality of the move and the current solution during a single point search.~\citeauthor{ref:burke:2003}~\cite{ref:burke:2003} mention that many move acceptance strategies have been explored within \acp{HH}.

Move acceptance strategies can be divided into two categories. These include \textit{deterministic} and \textit{non-deterministic} move acceptance strategies. Deterministic move acceptance strategies generate the same result for the same candidate solution(s) used for the move acceptance test. Non-deterministic move acceptance strategies can involve other parameters, such as the current time, or a sampling operation that yields possibly different outcomes when repeated for the move acceptance test~\cite{ref:burke:2003}.

\subsubsection{Construction vs. Perturbation}

Selection and generation \acp{HH} can be further classified in terms of \textit{construction} and \textit{perturbation} mechanisms. According to~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}, construction approaches build a solution incrementally. Starting with an empty solution, the goal is to intelligently select and use construction heuristics to gradually build a complete solution. As such, the \acs{HH} framework is provided with a set of pre-existing construction heuristics.

Perturbation  approaches refer to approaches that start with a complete solution, generated either randomly or using simple construction heuristics.  Perturbation heuristics try to iteratively improve the current solution. As such, the hyper-heuristic framework is provided with a set of neighbourhood structures and/or simple local searchers.

>>>>

~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} define \acp{HH} as search methods or learning mechanism for selecting or generating heuristics to solve computational search problems.~\citeauthor{ref:burke:2003}~\cite{ref:burke:2003} mentions that a \acs{HH} is a high-level heuristic approach that, given a particular problem instance and a number of low-level heuristics, can select and apply an appropriate low-level heuristic at each decision point. \acp{HH} implement a form of meta-learning that is concerned with the selection of the best heuristic from a pool of heuristics to solve a given problem. It can be said that \acp{HH} are concerned with finding the best heuristic in \textit{heuristic space}, while the underlying low-level heuristics find solutions in the feasible \textit{search/solution space}.

\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} proposed a modern classification scheme used to classify \acp{HH}. According to the proposed classification scheme, \acp{HH} are classified in two dimensions. These include the \textit{source of feedback} used during learning and the nature of the \textit{heuristic search space}. For the dimension that involves the source of feedback, \acp{HH} can be classified as either \textit{no learning}, \textit{online learning} or \textit{offline learning}. For the dimension that involves the nature of the \textit{heuristic search space}, \acp{HH} can be classified as either \textit{heuristic selection} or \textit{heuristic generation}. Further distinction is made between construction heuristics and perturbation heuristics.

In the general context of optimisation, many different types of \acp{HH} have been implemented and applied to many different problems. Some notable examples include \cite{ref:dowsland:2007, ref:burke:2010, ref:grobler:2012, ref:vanderstockt:2018}. Research on the application of \acp{HH} in the context of \acs{FFNN} training is still scarce.~\citeauthor{ref:nel:2021}~\cite{ref:nel:2021} provides some of the first research in this field, applying a \acs{HH} to \acs{FFNN} training.

This research takes a particular interest in a population-based, selection approach for \acp{HH}, with the particular intent of training \acp{FFNN}. In the context of population-based \acp{HH}, an entity pool exists that represent a pool of candidate solutions to the given problem. Each entity in the entity pool is assigned its own low-level heuristic from the heuristic pool. The selection of the best heuristic to apply to a candidate solution, is based on the performance of the heuristic relative to that particular candidate solution at a particular point in the search process. Selection methods often make use of probabilistic approaches.
